Sardines only end-to-end atlantisom test

### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
text edits
add survey and SS writing parts

@@

@@

## "Data" from OM for EMs

```{r survey-spec}

@@

Specify our survey sampling. This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc. Here we guess using Christine's specifications:

@@

#Set species name
species <- "Pacific_sardine" #change to species_ss to match

@@

@@

Get input survey biomass and catch biomass for SS (these have q, selectivity, cv)

```{r toSS-surveyfisheryts}

@@

@@

Get population parameters for SS

@@

>>>>

# save true catch at age
saveRDS(catch_numssshigh, file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))

@@

catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))

@@

@@

Here we are using Christine's specifications, which approximate the acoustic survey index used in the sardine assessment:

@@

#Set species name
species <- species_ss #"Pacific_sardine" #change to species_ss to match

@@

#Efficiency of the survey
effic <- data.frame(species=species, efficiency=eff_case)

# Assume uniform selectivity, same as selex1 here
sel<- data.frame(species=rep(species,length(age_classes)),
               agecl=age_classes,
              selex=sel_by_age)

@@

Get input survey biomass for SS (these have q, selectivity, cv)

```{r toSS-surveyts}

#Sample survey - 3rd timestep simulates a summer survey
survey_out <- create_survey(dat=truth$nums, 
                            time=survey_sample_full, 
                            species=species, 
                            boxes=boxall, 
                            effic=effic, 
                            selex=sel)

#Set effective sample size for age compositions
effN <- surveyEffN
highEffN <- data.frame(species=species, effN=rep(effN, length(species)))

#Sample fish for age composition
age_comp_data <- sample_fish(survey_out, highEffN)

# aggregate true resn per survey design
survey_aggresnstd <- aggregateDensityData(dat = results$resn,
                                          time = unique(results$nums$time),
                                          species = species,
                                          boxes = boxall)

# aggregate true structn per survey design
survey_aggstructnstd <- aggregateDensityData(dat = results$structn,
                                             time = unique(results$nums$time),
                                             species = species,
                                             boxes = boxall)

ss_structnstd <- sample_fish(survey_aggstructnstd,
                             effN,
                             sample=FALSE)
ss_resnstd <- sample_fish(survey_aggresnstd,
                          effN,
                          sample=FALSE)

#Extract length composition data
ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
                                     resn = ss_resnstd,
                                     nums = age_comp_data,
                                     biolprm = results$biolprm, fgs = results$fgs,
                                     CVlenage = CVs$lenage, remove.zeroes=TRUE)

#Need to replace with interp function
wtAtAge <- ss_length_stdsurv$muweight %>%
  select(species, agecl, time, wtAtAge = atoutput) %>%
  mutate(wtAtAge = wtAtAge/1000)

# CV function
cv <- data.frame(species="Pacific_sardine", cv=CVs$survey)

#Sample survey biomass
survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)

# survey numbers, not sure which SS needs?
survObsNum <- sample_survey_numbers(dat=survey_out,cv=cv)

Get input total catch for SS (here using numbers). SS takes a fishery cv which can be applied here to get a catch time series, or we can input true catch.

```{r toSS-fisheryts}

# catch at age by area
catch_numbers <-  create_fishery_subset(dat = truth$catch,
                                         time = fish_years,
                                         species = species,
                                         boxes = boxall)

# catch in numbers
catch_total <- catch_numbers %>%
  group_by(time) %>%
  summarise(catch=sum(atoutput))

# the above can be done with sample_survey_numbers and the fishery cv can be applied
# CV function
cv <- data.frame(species="Pacific_sardine", cv=CVs$fishery)

catchObsNum <- sample_survey_numbers(catch_numbers, cv=cv)

@@

#Set effective sample size for age compositions
effN <- surveyEffN
highEffN <- data.frame(species=species, effN=rep(effN, length(species)))

#Sample fish for age composition
age_comp_data <- sample_fish(survey_out, highEffN)

# Survey length comps and wtage done above to get survey ts using nums*wtage approach

#We end up using CAAL for the survey below, so let's generate fishery age comps instead
effN <- data.frame(species=species, effN=effN)

catch_numssshigh <- sample_fish(catch_numbers, effN)

#Get weights
# aggregate true resn per survey or fishery subset design
catch_aggresnss <- aggregateDensityData(dat = truth$resn,
                                 time = fish_years,
                                 species = species,
                                 boxes = boxes)

# aggregate true structn per survey or fishery subsetdesign
catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
                                 time = fish_years,
                                 species = species,
                                 boxes = boxes)

#dont sample these, just aggregate them using median
catch_structnss <- sample_fish(catch_aggstructnss, effN, sample = FALSE)

catch_resnss <-  sample_fish(catch_aggresnss, effN, sample = FALSE)

catch_length <- calc_age2length(structn = catch_structnss,
                                 resn = catch_resnss,
                                 nums = catch_numssshigh,
                                 biolprm = results$biolprm, fgs = results$fgs,
                                 maxbin = maxbin,
                                 CVlenage = CVs$lenage, remove.zeroes=TRUE)

@@

Get population parameters for SS from Atlantis

@@

## Write data for Stock Synthesis

To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`. You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.

```{r readdata, echo=FALSE}
require(r4ss)
#source("./config/sardine_config.R") this is above for illustration

stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
model_dir,
datfile_name))
```

Write time series biomass (tons) and catch (numbers) data for SS using `SS_write_ts`:

```{r SS-write-ts}
#Test writing CPUE in biomass
  stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
                ts_data = list(survObsBiom$atoutput[fish_years],
                  catch_total$catch[fish_years]),
                CVs = c(CVs$survey,
                        CVs$fishery),
                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             catch_total$time[fish_years]),
            sampling_month = list(rep(survey_month,nyears),
                                rep(fishing_month,nyears)),
                units = c("biomass","numbers"),
                data_type=c("CPUE","catch"),
                fleets = c(2,1))
```

Writing age composition data

```{r SS-write-comps}
#age_comp_data

#Get the age bins
age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
age_bins <- sub("a","",age_bin_names)

age_comp_flat <- reformat_compositions(age_comp_data,
                     round.places = 4,
                     comp_type = "agecomp")

## Write age composition data for survey
stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data,
               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
               sampling_month = list(rep(survey_month,nyears)),
               data_type = c("agecomp"),
               fleet_number = c(1),
               bins = list(age_bins),
               caal_bool = c(FALSE))
stocksynthesis.data$agecomp

#Check length compositions match age compositions
ss_length_stdsurv$natlength %>%
  group_by(time,agecl) %>%
  summarise(sum(atoutput))

len_comp_data <- ss_length_stdsurv$natlength
fish_len_comp_data <- catch_length$natlength

Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS

```{r SS-comps-formatting}

caal_comp_flat <- reformat_compositions(len_comp_data,                                round.places=4,
                comp_type="caalcomp")

#remove burnin
caal_comp_final <- filter(caal_comp_flat,
                         time %in% survey_years)

#Add over age classes to get sample size
len_comp_flat <- reformat_compositions(len_comp_data,
                                  round.places = 0,
                           comp_type="lencomp")
#remove burnin
len_comp_final <- filter(len_comp_flat,
                         time %in% survey_years)

length_bins <- as.integer(names(len_comp_final))
length_bins <- length_bins[!is.na(length_bins)]

fish_len_comp_flat <- reformat_compositions(catch_length$natlength,
                                  round.places = 0,
                           comp_type="lencomp")
#remove burnin
fish_len_comp_final <- filter(fish_len_comp_flat,
                         time %in% fish_years)

fish_age_comp_flat <- reformat_compositions(catch_numssshigh,
                           comp_type="agecomp")
#remove burnin
fish_age_comp_final <- filter(fish_age_comp_flat,
                         time %in% fish_years)

comp_list <- list(caal_comp_final,len_comp_final,          fish_age_comp_final, fish_len_comp_final)

apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), rep(survey_month, nrow(comp_list[[2]])), rep(fishing_month,nrow(comp_list[[3]])),rep(fishing_month,nrow(comp_list[[4]])))

# Write CAAL and length composition data
stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data, 
                                      comp_matrix = comp_list, 
                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
                                      sampling_month = apply_month, 
                                      data_type = rep(c("agecomp", "lencomp"),2), 
                                      fleet_number = c(2,2,1,1),  
                                      bins = list(age_bins, 
                                                  length_bins, 
                                                  age_bins, 
                                                  length_bins), 
                                      caal_bool = c(TRUE, rep(FALSE,3)))

head(stocksynthesis.data$lencomp)
head(stocksynthesis.data$agecomp)

#Change length bin structure to match atlantis data
stocksynthesis.data$lbin_vector <- length_bins

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)

SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
datfile_name),
                 overwrite=TRUE)
```

## Run Stock Synthesis
After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.

```{r runSS, eval=FALSE}
#Run stock synthesis model
run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
